Synergistic retrievals of ice in high clouds from elastic backscatter lidar, Ku-band radar and submillimeter wave radiometer observations

### 2023-11-09 RST/Submission/responseReviewers_Rev2.3.tex

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage[margin=1in]{geometry}
\author{Mircea Grecu and John E. Yorks}
\title{Synergistic retrievals of ice in high clouds\\
Response to Reviewers\\Second Revision}
\date{}
\begin{document}
\maketitle

\noindent \textbf{Reviewer 1}\\

\noindent\textit{The authors should state which version of the 2C-ICE and 2B-GEOPROF products they are using (R05?). Other than that, I think this is a good paper that is 
ready for publication.}\\

\noindent We thank the reviewer for the positive feedback. We added a sentence to the manuscript to clarify that we use the R05 version of the 2C-ICE and 2B-GEOPROF products.\\

\noindent \textbf{Reviewer 2}\\

\noindent\textit{Although I would prefer not to drag out this review process even longer, I feel that the authors failed to adequately address some of my previous comments. Most importantly, the published code does not contain the retrieval you describe in your paper. The repository only includes a retrieval based on a similarity search, which yields better results than the results presented in the manuscript. If these results are representative, they do put much of the presented results into question. I, therefore, suggest the authors provide the actual code used to produce the results presented in the manuscript instead.}\\

\noindent We thank the reviewer for the perspective.\\

\noindent \textbf{Other comments:}\\

\noindent\textit{\textbf{1.}Please note that the sentence above clearly states that $N_w$ depends on $a_{norm}$. I do understand that, in general, $N_w$ does not depend on $a_{norm}$. In your study, however, you are trying to infer both $N_w$ and $a_{norm}$ from the Z-IWC relationship in CS 2C-ICE. You describe how you use these values to derive the variation of $ln(a)$ and b, but that is still insufficient to determine $N_w$ and its altitude variation. Please provide the final parametrization of $N_w$ that you use in your simulations --prior to perturbing $N_w$ -- and how you calculated it. Currently, the manuscript only states that you calculate  Z for a fixed $N_w$, but it is unclear how you use that to determine $a_{norm}$.}\\

\noindent We are sorry about the confusion. Parameter $a_{norm}$ is derived using analytical normalized gamma distributions with $N_w$=0.08cm$^{-4}$. The use of analytical distributions in the derivation of the scattering lookup tables is already explained in the manuscript, but their relevance to $a_{norm}$ is insufficiently described. We added a few sentences to the manuscript to clarify this point.  Specifically, given that normalized particle size distributions are functions of type $N(D)=N_w\cdot F(D/D_m)$, normalization by $N_w$ transforms the relation between normalized $IWC$ and $Z$ to a relation between $IWC_{norm}=\int_0^{\infty} F(D/D_m)m(D)dD$ and $Z_{norm}=\frac {\lambda ^4} {\pi ^5 |K_w|^2} \int_0^{\infty} F(D/D_m) \sigma_b(D) dD$, where $m(D)$ and $\sigma_b(D)$ are the mass and electromagnetic backscattering cross-section of particles of mass $D$. As apparent in the above integrals, there is no $N_w$ dependence in the relations between normalized variables. A relationship of the type $IWC_{norm}=a_{norm}Z_{norm}^b$ becomes $IWC/N_w=a_{norm}(Z/N_w)^b$ for a PSD with a scaling intercept of $N_w$, or $IWC=a_{norm}N_w^{1-b}Z^b$. The values associated with our scattering lookup tables are $a_{norm}$ $\approx$0.985 and b$\approx$0.64, which result in parameterization $log_{10}(N_w)=8.87+0.161H$.\\

\noindent\textit{\textbf{2.}
$>$ l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.\\
You did not change anything in the manuscript.}\\

\noindent We apologize for the omission. We certainly intended to, but did not keep proper track of all changes meant to be included in the submission. We added a sentence to the manuscript to clarify this point.\\

\noindent\textit{Minor mistakes:\\
l. 94: Formatting of reference\\
l. 108: Space after full stop.\\
l. 121: Spelling mistake\\
l. 361: There $\->$ The\\
l. 441: 'Of' repeated\\
l. 462: Is $\->$ are.} \\

\noindent We thank the reviewer for pointing out these issues. We addressed them in the new version of the manuscript. It should be mentioned though that reference formatting is determined by the Latex compiler at submission time and some formatting aspects are not under the author's control.\\

\begin{thebibliography}{999}

\bibitem{arlot2010survey}
Arlot, S., A survey of cross-validation procedures for model selection. Statistics Surveys. 2010, 4, 40-79.

%\bibitem{ferreira2001}
%Ferreira, F., P. Amayenc, S. Oury, and J. Testud, 2001: Study and Tests of Improved Rain Estimates from the TRMM Precipitation Radar. J. Appl. Meteor. Climatol., 40, 1878–1899, https://doi.org/10.

\end{thebibliography}

\end{document}

### 2023-11-09 RST/Submission/RSTManuscript_rev3.tex
revision 3

% \subsection*{subsection}
% text...
\section{Methodology}
As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical distributions of ice properties needed in the investigation.  Although research quality CS cloud ice products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts and associated properties directly from CS reflectivity observations.  This ensures the consistency between the particle distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing and those used the simulation of the lidar, Ku-band radar and radiometer observations. Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.  When the CALIPSO lidar detects echo associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our CS-based estimates. Specifically, we use the IWC estimates from the 2C-ICE product in conjunction with a Particle Size Distribution (PSD) generalized intercept \citep{testud2001} provided by statistical model described in the next section to characterize the ice PSDs of clouds with radar echo below the noise level of the CloudSat radar. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated from CS observations using accurate physical models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm \cite{mackay2003information} is used to investigate the instrument synergy.  Details of the methodology are presented below.

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter, i.e. the normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001}, \cite{ferreira2001} and \cite{delanoe2014} showed that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}. A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power law ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals. The default multiplicative coefficient (the value that provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation with the relative height. Given that any deviation of the multiplicative coefficient in an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value \citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.

@@

Further insight into the radiometer-only estimation performance may be derived by defining the ice profile gravity center (GC) as $z_{GC}=\frac {\int_0^{Z_{top}}zIWC(z)dz} {\int_0^{Z_{top}}IWC(z)dz}$, where $z$ is the distance relative to the freezing level, the $Z_{top}$ is the distance from the top of the atmosphere to the freezing level. Shown in Figure \ref{f6} is the frequency of IWC gravity center estimated from radiometer-only observations as a function of its true value.  It may be observed in the figure that while the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only observations exhibits a multimodal narrow distribution. Moreover, there correlation between the retrieved and the true IWC gravity center is rather low.  This is another indication that, while the total amount of ice may be reasonably estimated from radiometer-only observations, its vertical distribution can not be accurately determined from radiometer-only observations.

@@

Additional insight may be gained by examining the distribution of the retrieved of IWP and IWC gravity center as a function of their true values.  These distributions are show in Figure \ref{f11} for the lidar-radiometer and radar-radiometer retrievals.  As apparent in the figure, the inclusion of radiometer observations improves both the lidar and radar retrievals. However, neither of the two instrument combinations is able to provide estimates that are as accurate as those derived from all instruments.  Moreover, the lidar-radiometer IWP retrievals appear to be less accurate than lidar-only retrievals. This may be an artifact of statistical inversion methodology, which may make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of profiles at the expense of others.  While other statistical inversion methodologies may be able to eliminate this artifact, given the complexity of the problem and the multiple sources of uncertainty, no better methodology is
obvious yet.

@@

\begin{table}[t]
\caption{Performance summary.}\label{t1}
\begin{center}
\begin{tabular}{c|ccccccc}
\hline\hline
\backslashbox{Score}{Instruments} & Lidar-& Radar- & Radiometer- & Lidar- & 
Lidar- & Radar-	& Lidar-\\
 & Only & Only & Only & Radar & Radiometer& Radiometer & Radar-Radiometer\\
\hline
% NRMS & 0.73 &  0.59  &  0.32 & 0.32 \\
NRMS & 0.84	& 0.67 & 0.64 & 0.61 &	0.56 &	0.54 &	0.48 \\
%0.73 &  0.59  &  0.32 & 0.22 \\
Class. Accurracy & 0.40	& 0.39 &	0.35 &	0.49 &	0.62& 0.53	& 0.64 \\
\hline
\end{tabular}
\end{center}
\end{table}

% text...
% 
\section{Conclusions}
In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities, etc.) are set based on the expected characteristics of instruments of the AOS mission. A hold-out validation methodology is used to assess the accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments. Specifically, the IWC and associated observations is randomly divided into two datasets, one for training and the other for evaluation.  The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DATA AVAILABILITY STATEMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
%
\datastatement
The CloudSat data can be accessed at: https://www.cloudsat.cira.colostate.edu/.
%  The data availability statement is where authors should describe how the data underlying 
%  the findings within the article can be accessed and reused. Authors should attempt to 
%  provide unrestricted access to all data and materials underlying reported findings. 
%  If data access is restricted, authors must mention this in the statement. See
%  {http://www.ametsoc.org/PubsDataPolicy} for more info.

>>>>

% \subsection*{subsection}
% text...
\section{Methodology}
As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical distributions of ice properties needed in the investigation.  Although research quality CS cloud ice products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts and associated properties directly from CS reflectivity observations.  This ensures the consistency between the particle distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing and those used the simulation of the lidar, Ku-band radar and radiometer observations. Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) version R05 of \cite{deng2015}.  When the CALIPSO lidar detects echo associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our CS-based estimates. Specifically, we use the IWC estimates from the 2C-ICE product in conjunction with a Particle Size Distribution (PSD) generalized intercept \citep{testud2001} provided by statistical model described in the next section to characterize the ice PSDs of clouds with radar echo below the noise level of the CloudSat radar. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated from CS observations using accurate physical models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm \cite{mackay2003information} is used to investigate the instrument synergy.  Details of the methodology are presented below.

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter, i.e. the normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001}, \cite{ferreira2001} and \cite{delanoe2014} showed that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs. Parameters $a_{norm}$ and $b$ are usually derived from regression of $IWC/N_w$ against $Z/N_w$ in the log-log space. Given that normalized PSDs are functions of type $N(D)=N_w\cdot F(D/D_m)$, normalization by $N_w$ removes any $N_w$ dependence from $IWC/N_w$ against $Z/N_w$ regressions, as both  $IWC/N_w$ and $Z/N_w$ are exclusive functions of $F(D/D_m)$. In this study, we derive $a_{norm}$ and $b$ from IWC and associated Z calculated from analytical normalized PSDs.  While details are present in subsequent paragraphs, it should be mentioned that the actual values of $a_{norm}$ and $b$ are 0.985 and 1.64, with $a_{norm}$ valid when the units of IWC, $N_w$ and $Z$ are g/m$^3$, m$^{-4}$, and dBZ respectively. However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature or height \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize $N_w$ as a function of height based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}. A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power law ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals. The default multiplicative coefficient (the value that provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level. Specifically, a regression of the type $\log_{10}(IWC)=\log_{10}(a)+b\cdot\log_{10}(Z)$ is performed for each bin referenced relatively to the zero degree bin is performed, and the variation of $\log_{10}(a)$ is parameterized as a function of the relative height $H$.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation with the relative height. Given that any deviation of the multiplicative coefficient in an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value \citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a function of the relative height.  Specifically, given that $(1-b)\log_{10}(N_w)+\log_{10}(a_{norm})=\log_{10}(a)$ and $\log_{10}(a)=\log_{10}(a_0)+s\cdot H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$. The actual expression derived in this study is $\log_{10}N_w=8.87+0.161\cdot H$ with $N_w$ expressed in m$^{-4}$ and $H$ expressed in km.

@@

Further insight into the radiometer-only estimation performance may be derived by defining the ice profile gravity center (GC) as $z_{GC}=\frac {\int_0^{Z_{top}}zIWC(z)dz} {\int_0^{Z_{top}}IWC(z)dz}$, where $z$ is the distance relative to the freezing level, the $Z_{top}$ is the distance from the top of the atmosphere to the freezing level. Shown in Figure \ref{f6} is the frequency of IWC gravity center estimated from radiometer-only observations as a function of its true value.  It may be observed in the figure that while the true IWC gravity center exhibits quite a broad distribution, the one retrieved from the radiometer-only observations is characterized by a narrower range with most values occurring between 2.0 km above the freezing level to about 5.5 km above the freezing level. Moreover, the correlation between the retrieved and the true IWC gravity center is rather low.  This is another indication that, while the total amount of ice may be reasonably estimated from radiometer-only observations, its vertical distribution can not be accurately determined from radiometer-only observations.

@@

Additional insight may be gained by examining the distribution of the retrieved IWP and IWC gravity center as a function of their true values.  These distributions are show in Figure \ref{f11} for the lidar-radiometer and radar-radiometer retrievals.  As apparent in the figure, the inclusion of radiometer observations improves both the lidar and radar retrievals. However, neither of the two instrument combinations is able to provide estimates that are as accurate as those derived from all instruments.  Moreover, the lidar-radiometer IWC gravity center retrievals appear to be less accurate than lidar-only retrievals. This may be an artifact of statistical inversion methodology, which may make suboptimal use of correlations in the observations and improve the accuracy of the estimates for some type of profiles at the expense of others.  While other statistical inversion methodologies may be able to eliminate this artifact, given the complexity of the problem and the multiple sources of uncertainty, no better methodology is
obvious yet.

@@

\begin{table}[t]
\caption{Performance summary.}\label{t1}
\begin{center}
\begin{tabular}{c|ccccccc}
\hline\hline
\backslashbox{Score}{Instruments} & Lidar-& Radar- & Radiometer- & Lidar- & 
Lidar- & Radar-	& Lidar-\\
 & Only & Only & Only & Radar & Radiometer& Radiometer & Radar-Radiometer\\
\hline
% NRMS & 0.73 &  0.59  &  0.32 & 0.32 \\
IWC NRMS & 0.84	& 0.67 & 0.64 & 0.61 &	0.56 &	0.54 &	0.48 \\
IWP NRMS & 1.12 & 0.54 & 0.4  & 0.48 &  0.37 &  0.32 &  0.3  \\
GC NRMS  & 0.21 & 0.58 & 0.46 &  0.26 &  0.23 &  0.42 &  0.19 \\
%0.73 &  0.59  &  0.32 & 0.22 \\
Class. Accurracy & 0.40	& 0.39 &	0.35 &	0.49 &	0.62& 0.53	& 0.64 \\
\hline
\end{tabular}
\end{center}
\end{table}

% text...
% 
\section{Conclusions}
In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities, etc.) are set based on the expected characteristics of instruments of the AOS mission. A hold-out validation methodology is used to assess the accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments. Specifically, the IWC and associated observations are randomly divided into two datasets, one for training and the other for evaluation.  The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DATA AVAILABILITY STATEMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 
%
\datastatement
The CloudSat 2C-ICE and 2B-GEOPROF version R05 data can be accessed at: https://www.cloudsat.cira.colostate.edu/. Code to simulate the synthetic observations from the CloudSat products and investigate the synergy of the instruments may be accessed at:  https://github.com/mirgrecu/synergIceRetrievals/.

%  The data availability statement is where authors should describe how the data underlying 
%  the findings within the article can be accessed and reused. Authors should attempt to 
%  provide unrestricted access to all data and materials underlying reported findings. 
%  If data access is restricted, authors must mention this in the statement. See
%  {http://www.ametsoc.org/PubsDataPolicy} for more info.
